Details of specific functions

### 2022-07-21 inst/gitbook/03_specific-tutorials.Rmd
vignette update

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

#### Projections and phylo projections: elaboration and exploration

@@

>>>>

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
1 | `disalignment` | The rejection of the centroid of a matrix from the major axis of another (typically an `"as.covar"` metric) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
1 | `projections.between` | `projections` metric applied between groups | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

#### Projections and phylo projections: elaboration and exploration {#projections}

@@

##### `projections.between` and `disalignment`

These two metrics are typically based on variance-covariance matrices from a `dispRity` object that has a `$covar` component (see more about that [here](#covar)).
Both are based on the `projections` metric and can take the same optional arguments (more info [here](#projections)).
The examples and explanations below are based on the default arguments but it is possible (and easy!) to change them.

We are going to use the `charadriiformes` example for both metrics (see more about that [here](#covar)).

```{r}
## Loading the charadriiformes data
data(charadriiformes)

## Creating the dispRity object (see the #covar section in the manual for more info)
my_covar <- MCMCglmm.subsets(n = 50,
                             data = charadriiformes$data,
                             posteriors = charadriiformes$posteriors,
                             group = MCMCglmm.levels(charadriiformes$posteriors)[1:4],
                             tree = charadriiformes$tree,
                             rename.groups = c(levels(charadriiformes$data$clade), "phylogeny"))
```

The first metric, `projections.between` projects the major axis of one group (`matrix`) onto the major axis of another one (`matrix2`).
For example we might want to know how some groups compare in terms of angle (orientation) to a base group:

```{r}
## Creating the list of groups to compare
comparisons_list <- list(c("gulls", "phylogeny"),
                         c("plovers", "phylogeny"),
                         c("sandpipers", "phylogeny"))

## Measuring the angles between each groups
## (note that we set the metric as.covar, more on that in the #covar section below)
groups_angles <- dispRity(data = my_covar,
                          metric = as.covar(projections.between),
                          between.groups = comparisons_list,
                          measure = "degree")
## And here are the angles in degrees:
summary(groups_angles)
```

The second metric, `disalignment` rejects the centroid of a group (`matrix`) onto the major axis of another one (`matrix2`).
This allows to measure wether the center of a group is aligned with the major axis of another.
A disalignement value of 0 means that the groups are aligned. A higher disalignment value means the groups are more and more disaligned.
We can use the same set of comparisons as in the `projections.between` examples to measure which group is most aligned (less disaligned) with the phylogenetic major axis:

```{r}
## Measuring the disalignement of each group
groups_alignement <- dispRity(data = my_covar,
                              metric = as.covar(disalignment),
                              between.groups = comparisons_list)
## And here are the groups alignment (0 = aligned)
summary(groups_alignement)
```
